# Reproducible Research: Peer Assessment 1

```{r}

```


## Loading and preprocessing the data

First of all I will load the data

```{r}
a<-read.csv("activity.csv")
b<-subset(a, a$steps!=is.na(a$steps))
b$date<-as.vector(b$date)
b$date<-as.Date(b$date, "%Y-%m-%d")
s<-aggregate(b$steps, list(b$date), sum)
colnames(s)<-c("date", "sum_per_day")
```

## What is mean total number of steps taken per day?

This is the total number of steps taken per day!

```{r}
plot(s$date, s$sum_per_day, type = "h", xlab = "days", ylab = "total number of steps", main = " Total number of steps taken each day")
```

Now here the calculation of *Mean* and *Median* per day

```{r}
m<-aggregate(b$steps, list(b$date), mean)
colnames(m)<-c("date", "mean_per_day")
me<-aggregate(b$steps, list(b$date), median)
colnames(me)<-c("date", "median")
m<-cbind(m, me$median)
colnames(m)<-c("date", "mean", "median")
m$date<-as.character(m$date)
```

The result in this Table:
```{r, results="asis"}
library(xtable)
mt<-xtable(m)
print(mt, type="html")
```

## What is the average daily activity pattern?

```{r}
d<-a
d$date<-as.vector(d$date)
d$date<-as.Date(d$date, "%Y-%m-%d")
er<-aggregate(d$steps, list(d$interval), mean, na.rm=TRUE)
colnames(er)<-c("interval", "mean_per_interval")
plot(er[,1], er[,2], type="l", xlab = "5-minute interval", ylab = " average number of steps taken", main = "Average daily activity pattern")
```

```{r}
z<-subset(er, er$mean_per_interval == max(er[,2]))
```

**`r z[,1]`**  is the 5-minute interval contains the maximum number of steps, on average across all the days in the dataset. The average of the steps is **`r z[,2]`**.


## Imputing missing values
```{r}
not_value<-a[is.na(a[,1]),]
i<-length(not_value[,1])
```
The total number of missing values in the dataset is: **`r i`**.

```{r}
t<-aggregate(not_value$interval, list(not_value$date), factor)
u1<-as.vector(t[,1])
rownames(t)<-u1
u2<-as.data.frame(t[,2])
rr<-length(u1)
rr2<-length(er[,1])
for (r in 1:rr){
        uu<-d[d$date == u1[r],]
        line<-NULL
        line<-rownames(uu)
        for (w in 1:rr2){
                d[line[w],1]<-er[w,2]                        
        }
}
```

Plot:


```{r}
s1<-aggregate(d$steps, list(d$date), sum) 
colnames(s1)<-c("date", "sum_per_day") 
plot(s1$date, s1$sum_per_day, type = "h", xlab = "days", ylab = "total number of steps", main = " Total number of steps taken each day")
```






```{r}
m1<-aggregate(d$steps, list(d$date), mean)
colnames(m1)<-c("date", "mean_per_day")
me1<-aggregate(d$steps, list(d$date), median)
colnames(me1)<-c("date", "median")
m1<-cbind(m1, me1$median)
colnames(m1)<-c("date", "new_mean", "new_median")
m1$date<-as.character(m1$date)
```

Table:
```{r, results="asis"}
library(xtable)
mt1<-xtable(m1)
print(mt1, type="html")
```


## Are there differences in activity patterns between weekdays and weekends?

```{r}
ddf<-d
ddf<-cbind(ddf,weekdays(d$date))
ddf[,5]<-NA
cnames<-c("steps","date","interval","weekdays", "weekend")
colnames(ddf)<-cnames
p<-length(ddf$weekdays)
for (t in 1:p){
        if ((ddf$weekdays[t]=="Friday")|(ddf$weekdays[t]=="Saturday")|(ddf$weekdays[t]=="Sunday")){
                ddf[t,5]<-TRUE
        }
        else{
                ddf[t,5]<-FALSE  
        }
}
mweeked <-subset(ddf, ddf$weekend == TRUE)
mw<-aggregate(mweeked$steps, list(mweeked$interval), mean)
colnames(mw)<-c("interval", "step_average")
mweekday <- subset(ddf, ddf$weekend == FALSE)
mwd<-aggregate(mweekday$steps, list(mweekday$interval), mean)
colnames(mwd)<-c("interval", "step_average")
average<-aggregate(ddf$steps, list(ddf$interval), mean)
colnames(average)<-c("interval", "step_average")
mw$diff_mean <- mw$step_average-average$step_average
mwd$diff_mean <- mwd$step_average-average$step_average
```

Panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
The number of steps are derived by (mean of weekend or weekday steps in i(th) interval - mean of all day steps in the i(th) interval)

```{r}
par(mfrow = c(2,1))
plot(mw$interval, mw$diff_mean, type = "l", xlab = "interval", ylab = "Number of steps", main="Weekend")
plot(mwd$interval, mwd$diff_mean, type = "l", xlab = "interval", ylab = "Number of steps", main="Weekday")
```

